### 2018-03-08 paper/dirlim.tex
first version of siamese

\subsection{Dimensionality Reduction by Learning an Invariant Mapping}

The problem of dimensionality reduction, mapping a high dimensional group of entities onto a low dimensional manifold, is an important problem in many areas of machine learning. Dimensionality Reduction by Learning an Invariant Mapping(DrLIM) is a technique for solving this problem by mapping points using a set of previously known matches and non-matches.\cite{Hasdell:Siamese} The goal of this technique is to minimize the distance between matched pairs and maximize the distance between unmatched pairs. This is accomplished using a network architecture know as a Siamese network. In this architecture there are two deep neural networks that share weights attached to a final output layer that determines the distance between the output of the two hidden networks. The network is trained by inputing known pairs and training the network to minimize or maximize the loss function.