### 2018-05-09 paper/Triplet_loss.tex
rough draft of paper is up

\subsection{Triplet Loss}
While the previos aproach maps 2 entities at a time using 2 networks that share weights, this aproach maps 3 entities at a time with 3 networks: the anchor the positive and the negative. By doing this we can at the same time maximise the distance between two unmatched pairs and minimize the distance between two matched pairs.\cite{schroff2015facenet} We do this by minimizing the following loss function: \[ L(\vec{X}_a, \vec{X}_n, \vec{X}_p) = |\:||\vec{X}_a - \vec{X}_p||^2 - ||\vec{X}_a - \vec{X}_n||^2 + \alpha|\] where \vec{X}_a, \vec{X}_n, \vec{X}_p are the embedded anchor, negative and positive vectors and \alpha is a constant margin. The key choice here is which vectors to include in the training set. If too many easy triplets are including it will slow down the learning since the constraint will be easily satisfied. If too many hard triplets are included the network will learn a strange function based on unusual cases.