### 2018-05-09 paper/Triplet_loss.tex
next draft of paper

\subsection{Triplet Loss}
While the previos aproach maps 2 entities at a time using 2 networks that share weights, this aproach maps 3 entities at a time with 3 networks: the anchor the positive and the negative. The loss function minimizes the difference between the anchor and the positive at the same time as it maximises the difference between the anchor and the negative. By doing this we can deal with an unmatched pair and a matched pair at the same time: the anchor is one name, the positive is a match for that name and the negative is another name which is not a match.\cite{schroff2015facenet} We accomplish this by minimizing the following loss function: \[ L(\vec{X}_a, \vec{X}_n, \vec{X}_p) = |\:||\vec{X}_a - \vec{X}_p||^2 - ||\vec{X}_a - \vec{X}_n||^2 + \alpha|\] where $\vec{X}_a$, $\vec{X}_n$, $\vec{X}_p$ are the embedded anchor, negative and positive vectors and $\alpha$ is a constant margin. The key choice here is which vectors to include in the training set. If too many easy triplets are including it will slow down the learning since the constraint will be easily satisfied. If too many hard triplets are included the network will learn a strange function based on unusual cases