### 2018-10-21 old/paper/previos_work.tex
reorgonized directory

\section{Related Works}
There have been a number of attempts to solve the fuzzy join problem. Many of the attempts use various string matching algorithms for example \cite{Wang:String}. The main issue with using string matching comparisons is that they do not work across a wide variety of entities. To our knowledge no serious machine learning approach has been used for this problem. The problem of blocking has also been dealt with using MapReduce \cite{Vernica:MapReduce}. This approach requires a lot of computing power and, since they use a string matching approach in the end, is hard to generalize across many types of entities. There has also been work done on a related yet different problem of fuzzy joining on whole rows \cite{he_ganjam_chu_2015}. We deal with the case where we only have one column of information.