### 2018-10-22 old/paper/toread.tex

[15] R. Vernica, M. J. Carey, and C. Li, “Efficient parallel set-
similarity joins using mapreduce,” in SIGMOD ’10.
[16] R. Baraglia, G. D. F. Morales, and C. Lucchese, “Document
similarity self-join with mapreduce,” in ICDM ’10.
[17] S. Chaudhuri, V. Ganti, and R. Kaushik, “A primitive operator
for similarity joins in data cleaning,” in ICDE ’06.
[18] C. Xiao, W. Wang, X. Lin, and J. X. Yu, “Efficient similarity
joins for near duplicate detection,” in WWW ’08, 2008.

http://ilpubs.stanford.edu:8090/1006/1/fuzzy-joins.pdf